---
title: "Getting started with DiceKriging"
author: "Victor Trappler"
date: "2024-04-03"
output: pdf_document
---

# Imports
```{r}
library(R6)
library(ggplot2)
library(purrr)
library(patchwork)
library(httpgd)
library(formatR)
library(ggforce)
```
# MOO class
In order to keep track of MOO problem.
```{r}
MultiObj <- R6Class("MultiObj", list(
  f = list(),
  dim_in = NULL,
  input_names = NULL,
  output_names = NULL,
  initialize = function(dim_in) {
    self$dim_in <- dim_in
    self$input_names <- c()
    for (i in 1:self$dim_in) {
      self$input_names <- c(self$input_names, paste0("x", i))
    }
  },
  add_objective = function(closure) {
    self$f[[length(self$f) + 1]] <- closure
  },
  dim_out = function() {
    self$output_names <- c()
    for (i in seq_along(self$f)) {
      self$output_names <- c(self$output_names, paste0("f", i))
    }
    return(length(self$f))
  },
  evaluate = function(input_df) {
    df_res <- input_df
    for (i in 1:self$dim_out()) {
      out <- unlist(pmap(input_df, (self$f[[i]])))
      df_res <- cbind(df_res, data.frame(out))
    }
    colnames(df_res) <- c(self$input_names, self$output_names)
    return(df_res)
  }
))


.f1 <- function(x1, x2) {
  b1 <- 15 * x1 - 5
  b2 <- 15 * x2
  lhs <- (b2 - (5.1 / (4 * pi^2)) * b1^2 + 5 * b1 / pi - 6)^2
  rhs <- 10 * ((1 - (1 / (8 * pi))) * cos(b1) + 1)
  return(lhs + rhs)
}
.f2 <- function(x1, x2) {
  b1 <- 15 * x1 - 5
  b2 <- 15 * x2
  lhs <- -sqrt((10.5 - b1) * (b1 + 5.5) * (b2 + 0.5))
  cen <- -(1 / 30) * (b2 - (5.1 / (4 * pi^2)) * b1^2 + -6)^2
  rhs <- -(1 / 3) * ((1 - 1 / (8 * pi)) * cos(b1) + 1)
  return(lhs + cen + rhs)
}
```
Define MultiObjective Branin-Hoo function
```{r}
branin <- MultiObj$new(dim_in = 2)
branin$add_objective(.f1)
branin$add_objective(.f2)
```
## First plots
```{r}
x_plot <- seq(0, 1, 0.05)
y_plot <- seq(0, 1, 0.05)
df_grid <- data.frame(expand.grid(x1 = x_plot, x2 = y_plot))
df_res <- branin$evaluate(df_grid)
fc1 <- ggplot() +
  geom_contour_filled(data = df_res, mapping = aes(x = x1, y = x2, z = f1)) +
  coord_fixed()
fc2 <- ggplot() +
  geom_contour_filled(data = df_res, mapping = aes(x = x1, y = x2, z = f2)) +
  coord_fixed()

fc1 + fc2 + plot_annotation("Contour plots of f_1 and f_2")
```

## Estimation of the Pareto front and set
The estimation depends on the points where the function is evaluated.
The *true* pareto front is defined  as the pareto front defined using the regular grid `df_grid`
```{r}
estimate_pareto_front <- function(df_res) {
  D <- df_res[order(df_res$f1, df_res$f2, decreasing = FALSE), ]
  front <- D[which(!duplicated(cummin(D$f2))), ]
  return(front)
}


# df_eval_grid <- branin$evaluate(df_grid)
df_eval_grid <- branin$evaluate(data.frame(x1 = runif(2000), x2 = runif(2000)))
pareto <- estimate_pareto_front(df_eval_grid)
pareto_front_true <- geom_step(data = pareto, mapping = aes(x = f1, y = f2), alpha = 1, colour = "red", linewidth = 1.3)
pareto_front <- ggplot() + pareto_front_true
pareto_set_true <- geom_point(data = pareto, mapping = aes(x = x1, y = x2), colour = "red")
pareto_set <- ggplot() + pareto_set_true
for (i in 1:50) {
  df_samples <- data.frame(x1 = runif(10), x2 = runif(10))
  df_res <- branin$evaluate(df_samples)
  pareto <- estimate_pareto_front(df_res)
  pareto_front <- pareto_front + geom_step(data = pareto, mapping = aes(x = f1, y = f2), alpha = 0.1)
  pareto_set <- pareto_set + geom_point(data = pareto, mapping = aes(x = x1, y = x2))
}
(pareto_front + ggtitle("Pareto Front")) +
  (pareto_set + ggtitle("Pareto set")) +
  plot_annotation(title = "Pareto front under sampling uncertainty")
```




## GP regression
We use `DiceKriging` for the GP regression
### Mean and std
```{r}
library(DiceKriging)
n_training <- 10
df_samples <- data.frame(x1 = runif(n_training), x2 = runif(n_training))
df_res <- branin$evaluate(df_samples)
model1 <- km(
  formula = ~1, design = df_samples, response = df_res$f1
)
model2 <- km(
  formula = ~1, design = df_samples, response = df_res$f2
)
prediction <- predict(model1, newdata = df_grid, type = "UK", checkNames = FALSE)
predict_df <- data.frame(
  x1 = df_grid$x1, x2 = df_grid$x2, pred = prediction$mean, pred_sd = (prediction$sd),
  lower = prediction$mean - prediction$sd, upper = prediction$mean + prediction$sd
)
k_fc <- ggplot() +
  geom_contour_filled(data = predict_df, mapping = aes(x = x1, y = x2, z = pred)) +
  geom_point(data = df_samples, mapping = aes(x = x1, y = x2, color = "Sampled points"), color = "red") +
  coord_fixed()
k_sd <- ggplot() +
  geom_contour_filled(data = predict_df, mapping = aes(x = x1, y = x2, z = pred_sd)) +
  geom_point(data = df_samples, mapping = aes(x = x1, y = x2, color = "Sampled points"), color = "red") +
  coord_fixed()

k_fc + k_sd + plot_annotation("GP mean and GP sd of f1")
```


### Conditional simulations

```{r}
df_simulated <- data.frame(df_grid,
  f1 = t(simulate(model1, newdata = df_grid, cond = TRUE)),
  f2 = t(simulate(model2, newdata = df_grid, cond = TRUE))
)
fc1_sim <- ggplot() +
  geom_contour_filled(data = df_simulated, mapping = aes(x = x1, y = x2, z = f1)) +
  geom_point(data = df_samples, mapping = aes(x = x1, y = x2, color = "Sampled points"), color = "red") +
  coord_fixed()
fc2_sim <- ggplot() +
  geom_contour_filled(data = df_simulated, mapping = aes(x = x1, y = x2, z = f2)) +
  coord_fixed()

fc1_sim + fc2_sim + plot_annotation("Contour plots of simulated f_1 and f_2")
```
```{r}
df_simulations <- data.frame(df_grid,
  f1 = double(length(df_grid$x1)),
  f2 = double(length(df_grid$x1)), r = factor(length(df_grid$x1))
)
for (i in 1:20) {
  df_simulations <- rbind(df_simulations, data.frame(df_grid,
    f1 = t(simulate(model1, newdata = df_grid, cond = TRUE)),
    f2 = t(simulate(model2, newdata = df_grid, cond = TRUE)),
    r = as.factor(i)
  ))
}
paretos <- do.call(rbind, by(df_simulations, df_simulations["r"], estimate_pareto_front))
paretos <- paretos[-c(1), ]
```

```{r}
plt_paretos <- ggplot() +
  pareto_front_true +
  geom_step(data = paretos, aes(x = f1, y = f2, group = r), alpha = 0.2)
plt_paretos + ggtitle("samples of Pareto front, under GP uncertainty, regular grid")
```

```{r}
is_dominated <- function(dom, x) {
  return((all(dom$f1 > x$f1 | dom$f2 > x$f2)))
}

test_plt <- ggplot()
splitted <- split(paretos, paretos["r"])
df_f <- expand.grid(f1 = seq(-100, 200, length.out = 60), f2 = seq(-35, 0, length.out = 50))
df_f[, "isdominated"] <- 0
for (j in seq_along(splitted)) {
  for (i in seq_len(nrow(df_f))) {
    df_f[i, "isdominated"] <- df_f[i, "isdominated"] + is_dominated(splitted[[j]], df_f[i, ]) / (1 + length(splitted))
  }
}
```
```{r}
brks <- c(0., 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1., 2)
q95 <- df_f[df_f["isdominated"] >= 0.95, ]
q50 <- df_f[df_f["isdominated"] >= 0.50, ]
q05 <- df_f[df_f["isdominated"] >= 0.05, ]
test_plot <- test_plt +
  geom_contour_filled(aes(f1, f2, z = as.numeric(isdominated)), data = df_f, alpha = 1, breaks = brks) +
  scale_fill_brewer(palette = "RdBu") +
  geom_mark_hull(data = q95, mapping = aes(f1, f2, color = "blue", ), expand = 0, radius = 0, colour = "blue") +
  geom_mark_hull(data = q50, mapping = aes(f1, f2, color = "black", ), expand = 0, radius = 0, colour = "black") +
  geom_mark_hull(data = q05, mapping = aes(f1, f2, color = "orange", ), expand = 0, radius = 0, colour = "orange") +
  pareto_front_true +
  ggtitle("Probability (wrt cond simulations) of being not dominated")

test_plot
```

```{r}
# pareto <- estimate_pareto_front(df_res_sim)
pareto_front <- ggplot() +
  pareto_front_true +
  geom_step(data = pareto, mapping = aes(x = f1, y = f2), alpha = 0.1)
pareto_set <- ggplot() +
  geom_point(data = pareto, mapping = aes(x = x1, y = x2))

n_training <- 50
for (i in 1:100) {
  df_samples <- data.frame(x1 = runif(n_training), x2 = runif(n_training))
  df_res_sim <- data.frame(df_samples,
    f1 = t(simulate(model1, newdata = df_samples, cond = TRUE)),
    f2 = t(simulate(model2, newdata = df_samples, cond = TRUE))
  )
  pareto <- estimate_pareto_front(df_res_sim)
  pareto_front <- pareto_front +
    geom_point(data = pareto, mapping = aes(x = f1, y = f2), alpha = 0.1)
  pareto_set <- pareto_set +
    geom_point(data = pareto, mapping = aes(x = x1, y = x2))
}
pareto_front + pareto_set
```
